---
title: "Plotting PHO84 reporter data for justifying time point selections"
author: Josh Ayelazuno, edited by Bin He
date: "2026-04-08 (updated `r Sys.time()`)"
output:
  html_notebook:
    toc: true
    toc_depth: 4
    code_folding: hide
---

Load libraries
```{r setup, message=FALSE}
require(tidyverse)
require(here)
```

Set themes for ggplot. If you prefer a different default plot setting, change here.
```{r}
old <- theme_set(theme_bw(base_size = 14))
```

Set working directory
```{r}
i_am("analysis/20260408-PHO84-reporter-analysis.Rmd")
```

## Goal
- Plot the %G0/G1 over a 6h time course for the four species grown with or without phosphate

## Data
Setting up species plotting orders and color
```{r}
species_color <- c(
  "S. cerevisiae" = "blue",
  "C. glabrata"   = "red",
  "K. lactis"     = "#f8c434",
  "C. albicans"   = "#74ac4c"
)
```

These data were collected by Chris and Varshu during 2022-2023. The values in the table are clearly background subtracted, log transformed MFIs. The raw FCS files and extracted MFI data need to be located and verified later.

Read data
```{r}
in_file = here("input", "20230907_PHO84-gfp_flow.csv")
dat <- read_csv(in_file, name_repair = "unique_quiet") |>
  # Pivot longer to reshape the data from wide to long format
  pivot_longer(
    cols = -`Time(hrs)`,
    names_to = "variable",
    values_to = "value"
  ) |>
  # remove PHO84p-EGFP as it's the same for all data
  mutate(variable = str_replace(variable, "PHO84p-EGFP", "")) |>
  # split the variable column into species and replicate
  separate_wider_delim(variable, 
                       names = c("Species", "Replicate"),
                       delim = "...") |>
  mutate(Species = fct(
    str_replace(Species, "\\.", "\\. "), levels = names(species_color)
  )) |>
  rename(Time = `Time(hrs)`, logMFI = value)
```
Write the reformatted data out for future use
```{r}
write_csv(dat, here("input", "20230907_PHO84-gfp_flow_tidy.csv"))
```

## Plotting

```{r}
# normalize the logMFI values by min-max scaling within each species
tmp <- dat %>%
  drop_na() %>% 
  group_by(Species) %>%
  mutate(n_logMFI = logMFI / mean(logMFI[Time >= 3])) %>% 
  ungroup()

tmp |>
  ggplot(aes(x = Time, y = n_logMFI, color = Species)) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(fun.data = "mean_cl_boot", 
               geom = "pointrange", size = 0.2, linewidth = 0.5) +
  #geom_point(size = 0.6, alpha = 0.6) +
  #facet_wrap(Treatment ~ Species) +
  #facet_wrap(~ Treatment, ncol = 2) +
  scale_color_manual(values = species_color) +
  labs(x = "Time (h)", 
       y = "normalized log<sub>10</sub>(PHO84p-GFP)",
       color = "Species") +
  xlim(0, 4) +
  theme(
    panel.border = element_rect(linewidth = 1.5),
    legend.position = c(0.75, 0.25),
    legend.text = element_text(face = 3),
    legend.background = element_rect(color = "gray50"),
    axis.title.y = ggtext::element_markdown()
  )
ggsave(here("output", "20260408-4sps-PHO84p-EGFP-time-course.png"),
       width = 4.5, height = 4.5)
```